Governing through patronage: the bargain for education in decentralized Brazil

### 2020-10-13 paper/paper.tex
incorporate new framing for intro

% what do I need to bring the audience onboard?
% emphasize the existence of discretion, turnover, political manipulation of public sresources.
% in particular the existence of power-sharing arrangements between elected officials that go beyond electoral concerns

Across the world, vulnerable populations rely on governments for access to basic public services such as education. In the developing world, the quality of public education remains low: functional illiteracy is widespread, and basic mathematical operations beyond the reach of many \citep{vegas_raising_2007}. The World Bank refers to this phenomenon as ``learning poverty'': stunted learning when access to a schooling exists.\footnote{For more information, see \hyperlink{here}{https://www.worldbank.org/en/topic/education/brief/learning-poverty}.} These outcomes result from political choices, but a growing body of literature recognizes the important role bureaucratic institutions mediating these outcomes \citep{pepinsky_bureaucracy_2017,finan_personnel_2015}.

Understanding how these political decisions cascade down the bureaucracy remains a challenge. Public services the end-result of a series of interlocking steps: from personnel decisions to service provision \citep{pepinsky_bureaucracy_2017}, each link in this institutional chain of command connects political principals to the child sitting in the classroom \citep{}.

@@

>>>>

% what do I need to bring the audience onboard?
% emphasize the existence of discretion, turnover, political manipulation of public resources.
% in particular the existence of power-sharing arrangements between elected officials that go beyond electoral concerns

Across the world, vulnerable populations rely on governments for access to basic public services such as education. In the developing world, the quality of public education remains low: functional illiteracy is widespread, and basic mathematical operations beyond the reach of many \citep{vegas_raising_2007}. The World Bank refers to this phenomenon as ``learning poverty'': stunted learning when access to a schooling exists.\footnote{For more information, see \hyperlink{here}{https://www.worldbank.org/en/topic/education/brief/learning-poverty}.} There is growing recognition of the important role bureaucratic institutions play in shaping these outcomes \citep{pepinsky_bureaucracy_2017,finan_personnel_2015}, with a renewed focus on how public sector reforms can create positive incentives for bureaucratic performance \citep{duflo_incentives_2012, dal2013strengthening, rasul2018management}.

Bureaucratic institutions, however, can only be fully understood in conjunction with their political principals. In the developing world, politicians often manipulate bureaucracies to further private agendas, with detrimental consequences for state capacity and public goods provision \citep{evans_bureaucracy_1999,kohli_state-directed_2004,centeno_unpacking_2017}. These political decisions cascade down through bureaucratic institutions and into public services \citep{bersch_state_2017}.

@@

Show you which mechanism operates to hurt state capacity and lead to lower educational outcomes: staff turnover.
